---
title: "Advanced R - Individual Assignment"
author: "Alan Carton"
date: "23 May 2019"
output:
  pdf_document: default
  html_document: null
  toc: yes
  toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(plyr)
library(dplyr)     # To compute the `union` of the levels.
library(png)       # To include images in this document.
library(knitr)     # To include images inline in this doc.
library(moments)   # Skewness
library(e1071)     # Alternative for Skewness
library(glmnet)    # Lasso
library(caret)     # To enable Lasso training with CV.
library(data.table) # for logging results
# library("doFuture")
library(xgboost)
library(rpart)
library(ridge)
#install.packages("spFSR")
library(FSelector)
library(spFSR)
library(mlr)

seed_val <- 3012
set.seed(seed_val)

# Setting a baseline before removing initial columns.
err_type_ <- "MAPE"

## Setting the working directory.
setwd("G:/My Drive/Documents/IE_Notes/R_Programming/T3Advanced/Assignment/Individual")

# R
starttime <- Sys.time()
head <- data.table(err_id=numeric(), model=character(), alpha=numeric(), rmse=numeric(), mae=numeric(), mape=numeric());

colResults <- head;
lassoResults <- data.table(err_id=numeric(), model=character(), alpha=numeric(), lambda=numeric(), rmse_train=numeric());

```

# Introduction

Objective: Predict House Prices
• Data: house_price_train.csv & house_price_test.csv
• Due date: May 23rd
• Submission:
  Printable Report (pdf/word/ppt)
  R codes, notebooks, all in a unique structure folder with relative paths
  Test Prediction (csv)

• Open scope on modeling and reporting (all optional):
  ↘ Methods to evaluate
  ↘ Metrics used
  ↘ Type of report
  ↘ Visualization
  ↘ Github link with project history

• Evaluation Criteria:
  ↘ Coding Quality
  ↘ Modeling Design
  ↘ Test Performance (MAPE)
  ↘ Reporting

## R Functions

```{r important_functions, message=FALSE, warning=FALSE}

print_time <- function(sec){
  print (format(sec, "%X"))
}

# Function to split a dataset into training and validation.

splitdf <- function(dataframe, seed=42) {
  # Function originally had 1/1.5 split for training.  I prefer 75/25 or 80/20 so changed to .8

  index <- 1:nrow(dataframe)
  trainindex <- sample(index, trunc(length(index)*.75))
  trainset <- dataframe[trainindex, ]
  testset <- dataframe[-trainindex, ]
  list(trainset=trainset,testset=testset)
}

# Returns the glm model so can be used in other situations.  I wanted to try a prediction using the model..
get_glm_model <- function(training_dataset, validation_dataset, title, method="glm"){
  train_control_config <- trainControl(method = "repeatedcv", 
                                       number = 5, 
                                       repeats = 1, #allowParallel =T,
                                       returnResamp = "all")
  
  this.model <- FALSE
  
  if (method=="glm"){
        # Fit a glm model to the input training data
    this.model <-caret::train(price ~ ., 
                              data = training_dataset, 
                              method = method, 
                              metric = err_type_,
                              preProc = c("center", "scale"),
                              na.action = na.omit,
                              trControl=train_control_config) 
  }
  else if (method=="rf"){
          # Fit a Random Forest model to the input training data
    train_control_config <- trainControl(method = "none", allowParallel =T)
    
    this.model <- caret::train(price ~ ., data = training,
                               method = "rf",
                               ntree = 1000,
                               trControl = train_control_config,
                               # num.threads = (availableCores()-1),
                               tuneGrid = data.frame(mtry = 6))
    
  }
  else if (method=="rpart"){
          # Fit a Random Forest model to the input training data
  
      train_control_config <- trainControl(method = "repeatedcv", 
                                       number = 5, 
                                       repeats = 1,
                                       returnResamp = "all")
    
    this.model <- caret::train(price ~ ., data = training,
                               method = "rf",
                               ntree = 1000,
                               trControl = train_control_config,
                               tuneGrid = data.frame(mtry = 6))
    
  }
  else if (method=="xgboost"){

    #formulaString <- as.formula(paste('price ~ ', paste(names(training), collapse="+")))

    train_control_config <- trainControl(method="repeatedcv", number=3, repeats=1)
    
    gb.tuneGrid <- expand.grid(
    		eta = c(0.3,0.4,0.5), 
    		nrounds = c(50,100,150), 
    		max_depth = 2:4, gamma = 0, 
    		colsample_bytree = 0.8, 
    		min_child_weight = 1, 
    		subsample=1)
    
    this.model <- caret::train(price ~ .,
    		data=training_dataset, 
    		method="xgbTree",
    		metric = err_type_,
    		trControl=train_control_config, 
    		#nthread=(availableCores()-1),
    		tuneGrid=gb.tuneGrid
        )
  }
  else{
        # Fit a glm model to the input training data
    this.model <-caret::train(price ~ ., 
                              data = training_dataset, 
                              method = method, 
                              metric = err_type_,
                              preProc = c("center", "scale"),
                              na.action = na.omit,
                              #num.threads = (availableCores()-1),
                              trControl=train_control_config) 
  }
 

  

  return(this.model)
}

# Setting base model.
lm_model<-0

mae<-function(real, predicted){
  return(mean(abs(real-predicted)))
}

mape<-function(real,predicted){
  return(mean(abs((real-predicted)/real)))
}

rmse<-function(real,predicted){
  return(sqrt(mean((real-predicted)^2)))
}

lm.model <- function(training_dataset, validation_dataset, title_) {
  # Create a training control configuration that applies a 5-fold cross validation
  
  this.model <- get_glm_model(training_dataset, validation_dataset, title_)
  
  # Prediction
  
  pred_ <- predict(this.model, validation_dataset)
  pred_[ is.na(pred_)] <- 0 # To avoid null predictions
  real_ <- validation_dataset$price
  
  # Error of the model
  rmse_ <- rmse(real_, pred_)
  mape_ <- mape(real_, pred_)
  
  # Error in terms of the mean deviation between the predicted value and the price of the houses
  mae_ <- mae(real_, pred_)
  
  lm_model<<-this.model
      
  colResults <<- rbind(colResults, data.table(err_id = nrow(colResults)+1, model = title_, alpha = 0, rmse = rmse_, mae=mae_, mape=mape_));
  
  # Plot the predicted values against the actual prices of the houses
  my_data <- as.data.frame(cbind(predicted=(exp(pred_) -1), observed=(exp(real_) -1)))
  ggplot(my_data, aes(predicted, observed)) +
    geom_point() + geom_smooth(method = "lm") +
    labs(x="Predicted") +
    ggtitle(ggtitle(paste(title_, 'MAPE: ', format(round(mape_, 4), nsmall=4), ' --> Price ERROR:', format(round(mae_, 0), nsmall=0), 
                          ' :: ', sep=''))) +  
    scale_x_continuous(labels = scales::comma) + 
    scale_y_continuous(labels = scales::comma)
  

}

run_model <- function(title, model_){
  
  model <- get_glm_model(training, validation, title, model_)
  model.model <- model

  pred_ <- predict(model, validation)
  real_ <- validation$price
  # ERROR of the model
  
  # Compute Metrics
  rmse_ <- rmse(real_, pred_)
  mape_ <- mape(real_, pred_)
  
  # Error in terms of the mean deviation between the predicted value and the price of the houses
  mae_ <- mae(train_ColToAnalyse, train_ColToAnalyse)
      
  colResults <<- rbind(colResults, data.table(err_id = nrow(colResults)+1, model = model_, alpha = alpha_, rmse = rmse_, mae=mae_, mape=mape_));
  
  model.rmse = rmse_
  model.price_error <- mae_
  model.mape <- mape_
  
  
  # Plot the predicted values against the actual prices of the houses
  my_data <- as.data.frame(cbind(predicted=(exp(pred_) -1), observed=(real_)))
  ggplot(my_data, aes(predicted, observed)) +
  geom_point() + geom_smooth(method = "lm") +
  labs(x="Predicted") +
  ggtitle(ggtitle(paste(title, ':: MAPE: ', format(round(mape_, 4), nsmall=4), ' --> Price ERROR:', format(round(mae_, 0), nsmall=0), 
  					  ' :: ', sep=''))) +  
  scale_x_continuous(labels = scales::comma) + 
  scale_y_continuous(labels = scales::comma)
  
  return(list(model=model, pred=pred_))
  
}
# Data_CleanupFunctions

# Functions for Listing columns for nulls from any given dataset
list_na_cols <- function(df){
  na.cols <- which(colSums(is.na(df)) > 0)
  print(paste('There are', length(na.cols), 'columns with missing values'))
  if(length(na.cols)>0){
  print(sort(colSums(sapply(df[na.cols], is.na)), decreasing = TRUE))
}

return(na.cols)
}

# Set Factor Values for Columns
clean_set_factor <- function(df, col_name, level_vals){

  cols <- colnames(df)==col_name
  df[,cols] <- factor(df[,cols], levels=level_vals)
  
  return(df)
}


```



# Data Reading and preparation
The dataset is offered in two separated files, one for the training and another one for the test set.  They need to be merged for cleaning and data treatment, and split at a later date back to the original files. 

```{r data_reading, message=FALSE, warning=FALSE}

# First read the individual datasets into the code
original_training_data = read.csv(file = file.path("house_price_train.csv"))
original_test_data = read.csv(file = file.path("house_price_test.csv"))

# Datasets merged for cleaning purposes, will be split again post cleaning for train / validation / test.
original_test_data["price"] <- 0
dataset <- rbind(original_training_data, original_test_data)

na.cols <- list_na_cols(dataset)
```

To avoid applying the Feature Engineering process two times (once for training and once for test), you can just join both datasets (using the `rbind` function), apply your FE and then split the datasets again. However, if we try to do join the two dataframes as they are, we will get an error because they do not have the same columns: `test_data` does not have a column `price`. Therefore, we first create this column in the test set and then we join the data

Let's now visualize the dataset to see where to begin

```{r dataset_visualization, message=FALSE, warning=FALSE}

#
# id - identifier. Qty Records 217277
# Consisted of 21 Columns before cleanup and feature generation.

# For the purposes of visualization, Summary and head were used to identify the data and patterns.  No other visualization was used other than boxplots for outliers, which appear later.

# summary(dataset) # Commented out as takes lot of screenspace.
head(dataset)


```

We can see some problems just by taking a look to the summary: the dataset has missing values, there are some categorical columns codified as numeric, it has different scales for the feature values. In addition, I will recommend you to take a deeper look to the data to detect more subtle issues: correlation between features, skewness in the feature values...

# Data Cleaning

The definition of "meaningless" depends on your data and your intuition. A feature can lack any importance because you know for sure that it does not going to have any impact in the final prediction (e.g., the id of the house). In addition, there are features that could be relevant but present wrong, empty or incomplete values (this is typical when there has been a problem in the data gathering process). For example, the feature `Utilities` present a unique value, consequently it is not going to offer any advantage for prediction.

We remove meaningless features and incomplete cases.

```{r eliminate_unique_data, message=FALSE, warning=FALSE}

# Utilities and especially id are deemed to be meaningless, as both are unique to the records.

# id is a unique column so will add nothing further to the data.
dataset <- dataset[,-which(names(dataset) == "id")]
dataset <- dataset[,-which(names(dataset) == "date")]

# Some additional columns will be removed later as may be used for interpretation before removing.

```

## Hunting NAs

Our dataset is filled with missing values, therefore, before we can build any predictive model we'll clean our data by filling in all NA's with more appropriate values.
As another option, we could just remove the entries with null values (i.e., remove rows). However, in this situation (and in many other that you will face) this is out of the question: we have to provide a prediction for each and every one of the houses (required by the competition). 
Similarly, you could discard the features with null values (i.e., remove columns), but it would mean the removal of many features (and the information they provide).

As a rule of thumb, if you are allowed to discard some of your data and you do not have many null values (or you do not have a clear idea of how to impute them) you can safely delete them. If this is not the case, you must find a way to impute them (either by applying some knowledge of the addressed domain or by using some more advanced imputation method: https://topepo.github.io/caret/pre-processing.html#imputation)

Counting columns with null values. - 34 Columns have NA pre cleanup.

```{r , message=FALSE, warning=FALSE}

# 0 Columns have NA Values.
list_na_cols(dataset)

```

## How to clean up NAs

Assign them default values, and assign features the correct type?

In any case, what we do here, is simply go through every single **factor** feature to: extend the number of possible levels to the new default for NAs (`None` or  `No` for categorical features or any other default value described in the documentation). For numerical values, we can just change the NA value for a default value, the median of the other values or some other value that you can infer (i.e., for the null values in the `GarageYrBlt` column you could use the year the house was built as a replacement value).

```{r replace_na_by_value, message=FALSE, warning=FALSE}

#function to count the number of nulls in a given dataset column
are_nulls <- function(x_name){
  return(sum(is.na(x_name)) > 0)
}

# No other columns suggested replacements other than NA or None / 0

```

and replacing the potential NA values in the dataset by that new 'factor' level.
Similarly, for numerical values:

```{r numbers_as_Categorical, message=FALSE, warning=FALSE}
# Some numerical values need to be treated as categorical, and need to be replaced with factor levels.

filterCols <- c("bedrooms", "bathrooms", "floors", "waterfront", "view","condition", "grade", "yr_built", "yr_renovated", "zipcode")

# Some numeric fields should be factors (i.e. Number of Rooms, etc.)
# These do not have factors, but should be set to the existing values.
for(col_name in filterCols){
  level_vals <- sort(unique(dataset[col_name])[,col_name])
  dataset[,col_name] <- factor(dataset[,col_name], levels=level_vals)
  
}

na.cols <- list_na_cols(dataset)

```

## Factorize features

If we go back to the summary of the dataset we can identify some numerical features that are actually categories: `MSSubClass` and the Year and Month in which the house was sold. What we have to do is to convert them to the proper 'class' or 'type' using the `as.factor` command.

Also setting some variables which do not make sense to be treated as numerical.  Number of bedrooms, you would not be meaningful for an average of 2.3 bedrooms ... etc.

```{r factorize_features, message=FALSE, warning=FALSE}


# Calculating some new features here as required earlier in the process.

dataset["TotalFlArea"] <- dataset$sqft_above + dataset$sqft_basement

dataset["sqft_building15"]<- dataset$sqft_living15 + dataset$sqft_lot15
dataset["sqft_building"]<-dataset$sqft_living + dataset$sqft_lot

dataset["has_basement"]<-dataset$sqft_basement>0
dataset["increased15"]<-(dataset$sqft_building15 - dataset$sqft_building)>0
dataset["unchanged15"]<-(dataset$sqft_building15 - dataset$sqft_building)==0

filterCols <- c("increased15", "unchanged15", "has_basement")

# Some numeric fields should be factors (i.e. Number of Rooms, etc.)
# These do not have factors, but should be set to the existing values.
for(col_name in filterCols){
  level_vals <- sort(unique(dataset[col_name])[,col_name])
  dataset[,col_name] <- factor(dataset[,col_name], levels=level_vals)
  
}

```

## Outliers
We will now focus on numerical values. If `NAs` where the natural enemy of categorical values, the main problem with numerical values are outliers (values which largely differ from the rest). Outliers can mislead the training of our models resulting in less accurate models and ultimately worse results.

In this section we seek to identify outliers to then properly deal with them. If we summarize the dataset, we can see variables which "Max." is much larger than the rest of values. These features are susceptible of containing outliers. Nevetheless, the easiest way to detect outliers is visualizing the numerical values; for instance, by `boxploting` the column values.

*Tip:* As explained in the feature engineering practice, the `boxplot` function can eliminate the outliers. However, if you apply it with the default values it is going to eliminate too much of them. You can adapt its working with the `outlier.size` param (https://ggplot2.tidyverse.org/reference/geom_boxplot.html), which I recommend you to set to at least 3. Another thing you can do is to compare the columns with outliers to the target variable (`price`) to visually check if there are some extreme values and just consider those as outliers.

## Dealing with Outliers
When dealing with Outliers, Looked at removing the outliers, but so many values were to be removed that it did not make sense to do so.  Per the instructions above, with a minimum coef set to 3.  When looking at this, it was not possible to remove the outliers, as it could also have been removing values from the testing dataset as well.  

Although they could possibly have been ignored, I decided to set all outliers to their respective minimum outlier value.  They would still be showing as outlier, but within an approved magnitude.  Setting to 0 or a Mean / Median did not appear to me to be a suitable solution in this instance.  By setting to the minimum value of the outlier range, it both classes them as outliers, but minimising their effect on the dataset.

```{r outlier_detection, message=FALSE, warning=FALSE}

library(outliers)

# Function to retrieve list of numeric columns from a dataset, with option to include / exclude the target_column for models.
get_numeric_cols <- function(df_, target_column){
  ret_val <- colnames(df_[, sapply(df_, is.numeric) & colnames(df_) != target_column])
  return(df_[ret_val])
}

numeric_cols <- colnames(get_numeric_cols(dataset, 'price'))

# Using the boxplot stats function to retrieve the list of outlier values.
# Not possible to remove outliers, as would effect the dataset.
# By setting the values to the minimum outlier value accepted, it is still identifying them as outliers but minimising their effect on the dataset.

for(col_name in numeric_cols){
  coef_val <- boxplot.stats(dataset[,col_name], coef = 3)$out
  min_outlier <- min(coef_val, na.rm=FALSE)
  # locate the outlier records, and set their value to the min
  dataset[dataset[,col_name]  %in% coef_val , col_name] <- min_outlier

}

```

## Now Show Boxplots for outlier columns.  Will still show as outliers, but within approved distance.

Boxplots were originally shown for in one graphic for all columns, but the scale was very bad for some, so I decided to group into 4 different boxplot graphs.

```{r outliers_boxplots, message=FALSE, warning=FALSE}
# Boxplots were originally shown for all columns, but the scale was very bad for some, so I decided to group into 4 different boxplot graphs.

ShowBoxplot <- function(df, col_names) {
  boxplot(df[,col_names], xaxt="n", coef = 3, outlier.shape=NA, outline=FALSE)
  
  # x axis with ticks but without labels
  # Plot x labs at default x position
  text(x =  seq_along(col_names), y = par("usr")[3] - (par("usr")[4] - par("usr")[3])/30, srt = 45, adj = 1,
  labels = colnames(dataset[col_names]), xpd = TRUE)
}

# Basic box plot
ShowBoxplot(dataset, numeric_cols)

```


## Skewness

We now need to detect skewness in the Target value. Let's see what is the effect of skewness on a variable, and plot it using ggplot. The way of getting rid of the skewness is to use the `log` (or the `log1p`) of the values of that feature, to flatten it. To reduce right skewness, take roots or logarithms or reciprocals (x to 1/x). This is the commonest problem in practice. To reduce left skewness, take squares or cubes or higher powers.

```{r check_skewness, message=FALSE, warning=FALSE}

df <- rbind(data.frame(version="price",x=original_training_data$price),
            data.frame(version="log(price+1)",x=log(original_training_data$price + 1)))

ggplot(data=df) +
  facet_wrap(~version,ncol=2,scales="free_x") +
  geom_histogram(aes(x=x), bins = 50)


```

We therefore transform the target value applying log.

The same "skewness" observed in the target variable also affects other variables. To facilitate the application of the regression model we are going to also eliminate this skewness. For numeric feature with excessive skewness, perform log transformation

Setting threshold for the skewness in 0.75. 


```{r set_skew_threshold, message=FALSE, warning=FALSE}

# Log transform the target for official scoring
dataset$price <- log1p(dataset$price)

skewness_threshold = 0.75

```

Now, let's compute the skewness of each feature that is not 'factor' nor 'character'. So, I'm only interested in continuous values. One possible way of doing it is the following: First, lets determine what is the 'class' or data type of each of my features.

To do so, instead of `loops`, we will use the `apply` family of functions. They will __apply__ a method to each **row** or **column** of your dataset. It will depend on what to do specify as the first argument of the method. 

What we want to determine is the class of each column or feature, and to do so, we use the `class` method from R. We will pass the actual column or feature from our dataset (dataframe).  With that information, we need to calculate the skewness of each column whose name is our list of __factor__ (or categorical) features. We use the `sapply` method again, to compute the skewness of each column whose name is in the list of `numeric_columns`.

```{r apply_skewness_nonfactors, message=FALSE, warning=FALSE}
column_types <- sapply(names(dataset), function(x) {
  class(dataset[[x]])
}
)

# Getting the list of columns that are not factors
numeric_columns <- names(column_types[column_types != "factor"])

# Generalising the above in a function to only return a type or all except a type.
get_columns_of_type <- function(data_, col_type){
  col_types <- sapply(names(data_), function(x){class(data_[[x]])});
  ret_names <- names(col_types[col_types == col_type])
  return(ret_names)
}

get_columns_not_of_type <- function(data_, col_type){
  col_types <- sapply(names(data_), function(x){class(data_[[x]])});
  ret_names <- names(col_types[col_types != col_type])
  return(ret_names)
}

numeric_cols = numeric_cols[numeric_cols!="lat"]
numeric_cols = numeric_cols[numeric_cols!="long"]

skew <- sapply(numeric_cols, function(x) {e1071::skewness(dataset[[x]], na.rm = T)})

# transform all variables above a threshold skewness.
skew <- skew[abs(skew) > skewness_threshold]
for(x in names(skew)) {
  dataset[[x]] <- log(dataset[[x]] + 1)
}

```

What we do need to make now is to apply the log to those whose skewness value is below a given threshold that we've set in 0.75. We should test different hypothesis with our threshold too.

```{r apply_new_Skew_threshold, message=FALSE, warning=FALSE}

# transform all variables above a threshold skewness.
skew <- skew[abs(skew) > skewness_threshold]
for(x in names(skew)) {
  dataset[[x]] <- log(dataset[[x]] + 1)
}

```

# Feature Creation
This is the section to give free rein to your imagination and create all the features that might improve the final result. Do not worry if you add some "uninformative" feature because it will be removed by the later feature selection process.
Do not hesitate to consult the competition kernels (please cite anything you fork).



```{r create_new_features, message=FALSE, warning=FALSE}
# transform all variables above a threshold skewness.

pre_transform_df <- dataset
dataset <- pre_transform_df


cols_to_remove <- c("")

features_to_keep <- colnames(dataset[, !colnames(dataset) %in% cols_to_remove])

```



```{r train_test_val_split, message=FALSE, warning=FALSE}

len_training = NROW(original_training_data)
len_test = NROW(original_test_data)
len_all = NROW(dataset)

training_data <- head(dataset, n=len_training)
test <- tail(dataset, n=len_test)

splits <- splitdf(training_data, seed=1)
training <- splits$trainset
validation <- splits$testset

tmp_test <- test
tmp_validation <- validation
tmp_training <- training

# Additional Functions
reset_data <- function(){
  
  training <<- tmp_training;
  test <<- tmp_test;
  validation <<- tmp_validation;
  
  colResults
  
}

get_best_model <- function(err_type){

  val_ <- 0
  err_type = tolower(err_type)
  if(err_type=="mape"){
    val_ <- colResults[colResults$mape==min(colResults$mape)]
  }
  if(err_type=="mae"){
    val_ <- colResults[colResults$mae==min(colResults$mae)]
  }
    if(err_type=="rmse"){
    val_ <- colResults[colResults$rmse==min(colResults$rmse)]
    }
  # Return sorted by descending id.
  return(val_[,,-err_id])

}

set_backup <- function(model_){
  
  model_run <<- model_;
  tmp_test <<- test;
  tmp_validation <<- validation;
  tmp_training <<- training;
}

check_best_model <- function(err_type){
  best_ <- get_best_model(err_type)
  if (best_$model!=model_run){
    print("Model error deteriorated, reseting to last set of columns.")
    reset_data();
    }
}

```



```{r baseline_1, message=FALSE, warning=FALSE}

set_backup("Baseline-1")
# Using all columns

glmmodel <- lm.model(training, validation, model_run)
glmmodel

check_best_model(err_type_)

# Checking the Error values for comparison about whether or not to retain or remove the column suggestions.
print(colResults)

```

# Train, Validation Spliting

To facilitate the data cleaning and feature engineering we merged train and test datasets. We now split them again to create our final model.

We are going to split the modified dataset into Test, train and validation for the later evaluation of our regression models

```{r baseline_2_removal, message=FALSE, warning=FALSE}

```

# Feature Selection
We here start the Feature Selection.

## Filtering Methods
We will rank the features according to their predictive power according to the methodologies seen in class: the Chi Squared Independence test and the Information Gain.


#### Full Model

Let's try first a baseline including all the features to evaluate the impact of the feature engineering.

### Chi-squared Selection
let's use the chisq.test included in the base package of R, to measure the relationship between the categorical features and the output. Only those.


```{r baseline_1_chisquare, message=FALSE, warning=FALSE}
# Compute the ChiSquared Statistic over the factor features ONLY
features <- colnames(training[, sapply(training, is.factor) & colnames(training) != 'price'])

chisquared <- data.frame(features, statistic = sapply(features, function(x) {
  chisq.test(training$price, training[[x]])$statistic
}))

# Plot the result, and remove those below the 1st IQR (inter-quartile-range) --aggressive
par(mfrow=c(1,2))
boxplot(chisquared$statistic)
bp.stats <- as.integer(boxplot.stats(chisquared$statistic)$stats)   # Get the statistics from the boxplot

chisquared.threshold = bp.stats[2]  # This element represent the 1st quartile.
text(y = bp.stats, labels = bp.stats, x = 1.3, cex=0.7)
barplot(sort(chisquared$statistic), names.arg = chisquared$features, cex.names = 0.6, las=2, horiz = T)
abline(v=chisquared.threshold, col='red')  # Draw a red line over the 1st IQR

```



Now, we can test if this a good move, by removing any feature with a Chi Squared test statistic against the output below the 1 IQR.

```{r baseline_2, message=FALSE, warning=FALSE}

# Determine what features to remove from the training set.
features_to_remove <- as.character(chisquared[chisquared$statistic < chisquared.threshold, "features"])
col_names_remaining <- !colnames(training) %in% features_to_remove

print("Columns marked for potential removal:")
print(features_to_remove)

set_backup("baseline-2")

training <- training[col_names_remaining]
validation <- validation[col_names_remaining]
test <- test[col_names_remaining]

glmmodel <- lm.model(training, validation, model_run)
glmmodel

# Depending on results, either keep or retain the columns identified.

check_best_model(err_type_)
# Checking the Error values for comparison about whether or not to retain or remove the column suggestions.

```

```{r baseline_2_results, message=FALSE, warning=FALSE}
# Checking the Error values for comparison about whether or not to retain or remove the column suggestions.
print(colResults)
```


### Now, Try feature selection of numerical variables with Spearman's correlation.

What to do with the numerical variables? We can always measure its relation with the outcome through the Spearman's correlation coefficient, and remove those with a lower value. Let's repeat the same process we did with the Chi Square but modifying our code to solely select numerical features and measuring Spearman'.

```{r baseline_3_chisquare, message=FALSE, warning=FALSE}
# Compute the ChiSquared Statistic over the factor features ONLY
features <- colnames(training[, sapply(training, is.numeric) & colnames(training) != 'price'])

spearman <- data.frame(features, statistic = sapply(features, function(x) {
cor(training$price, training[[x]], method='spearman')
}))

# Plot the result, and remove those below the 1st IQR (inter-quartile-range) --aggressive
par(mfrow=c(1,2))
boxplot(abs(spearman$statistic))
bp.stats <- boxplot.stats(abs(spearman$statistic))$stats   # Get the statistics from the boxplot
text(y = bp.stats, 
labels = sapply(bp.stats, function(x){format(round(x, 3), nsmall=3)}), # This is to reduce the nr of decimals
x = 1.3, cex=0.7)

spearman.threshold = bp.stats[2]  # This element represent the 1st quartile.

barplot(sort(abs(spearman$statistic)), names.arg = spearman$features, cex.names = 0.6, las=2, horiz = T)
abline(v=spearman.threshold, col='red')  # Draw a red line over the 1st IQR

```

Let's train the model with the new features, exactly as we did in the Chi Sq. section above.

```{r baseline_3, message=FALSE, warning=FALSE}
# Determine what features to remove from the training set.
features_to_remove <- as.character(spearman[spearman$statistic < spearman.threshold, "features"])
col_names_remaining <- !colnames(training) %in% features_to_remove

print("Columns marked for potential removal:")
print(features_to_remove)
print("Columns marked to use:")
print(col_names_remaining)

set_backup("Baseline-3")

training <- training[col_names_remaining]
validation <- validation[col_names_remaining]
test <- test[col_names_remaining]

glmmodel <- lm.model(training, validation, model_run)
glmmodel

# Based on the results either keep or discard the above features.
check_best_model(err_type_)

# [1] "Columns marked for potential removal:"
# [1] "sqft_lot"   "long"       "sqft_lot15"

# Checking the MAPE values for comparison about whether or not to retain or remove the column suggestions.
# "Model error deteriorated, reseting to last set of columns."
```

```{r baseline_3_results, message=FALSE, warning=FALSE}
# Checking the Error values for comparison about whether or not to retain or remove the column suggestions.
print(colResults)
```


### Information Gain Selection

This part is equivalent to the Chi Squared, but with another metric. So, the coding is very much equivalent.

```{r baseline_3_information.gain, message=FALSE, warning=FALSE}

left <- training$price

# This computes the information gain and stores in Dataset
weights <- data.frame(information.gain(left ~ .,training))
weights["col_names"] <- rownames(weights)

# Recommended features to keep.
features_to_use <- weights$col_names[weights$attr_importance >= 0.05]

print("Columns marked to use:")
print(features_to_use)

#  [1] "Columns marked to use:"
#  [1] "price"           "bedrooms"        "bathrooms"       "sqft_living"     "floors"         
#  [6] "grade"           "sqft_above"      "sqft_basement"   "yr_built"        "zipcode"        
# [11] "lat"             "long"            "sqft_living15"   "sqft_lot15"      "TotalFlArea"    
# [16] "sqft_building15" "sqft_building"  
```


```{r baseline_4_check_checkresults, message=FALSE, warning=FALSE}

set_backup("Baseline-4")

training <- training[features_to_use]
validation <- validation[features_to_use]
test <- test[features_to_use]

glmmodel <- lm.model(training, validation, model_run)
glmmodel

# Based on the above, either keep or discard the columns identified
check_best_model(err_type_)

```
### Check Error values to see if columns can be removed

The removal of columns recommended from information gain checks did not result in improvement of rmse score, so they were retained,

```{r baseline_4_results, message=FALSE, warning=FALSE}
# Checking the error values for comparison about whether or not to retain or remove the column suggestions.
print(colResults)
```

```{r final.Baseline.models, message=FALSE, warning=FALSE}

bin_price <- function(x, col_, bins){
  var_retd <- min(which(x[col_] <= bins))
  return(var_retd)
}

#> Loading required package: ParamHelpers
target_column <- "price"
retain_columns <- colnames(training)
knnWrapper    <- makeLearner("classif.knn", k = 5)

# Just get the numeric columns from the dataset to work out which better predictors.
tmpset <- get_numeric_cols(training, '')
col_bins <- c(1:max(tmpset[target_column] + 1))

tmpset["new_factor"] <- apply(tmpset, 1, function(x) min(which(x[target_column] <= col_bins)))

tmpset <- clean_set_factor(tmpset, target_column, c(unique(tmpset[target_column])))
tmpset <- tmpset[colnames(tmpset) != target_column]

classifTask   <- makeClassifTask(data = tmpset, target = "new_factor")
perf.measure  <- acc


spsaMod <- spFeatureSelection(
  task = classifTask,
  wrapper = knnWrapper,
  measure = perf.measure ,
  num.features.selected = 3,
  iters.max = 10,
  num.cores = 1)

cols_to_remove <- colnames(tmpset[!colnames(tmpset) %in% c(spsaMod$features, "new_factor")])
cols_to_retain <- colnames(training[!colnames(training) %in% cols_to_remove])

print("Columns marked for potential removal:")
print(cols_to_remove)
print("Columns marked to use:")
print(cols_to_retain)

# [1] "Columns marked for potential removal:"
# [1] "sqft_living"     "sqft_above"      "sqft_basement"   "sqft_living15"   "sqft_lot15"     
# [6] "sqft_building15" "sqft_building" 

# [1] "Columns marked to use:"
# [1] "price"       "bedrooms"    "bathrooms"   "floors"      "grade"       "yr_built"    "zipcode"    
# [8] "lat"         "long"        "TotalFlArea"

```


```{r baseline_5_removal, message=FALSE, warning=FALSE}

glmmodel <- lm.model(training, validation, "Baseline-5")
glmmodel

set_backup("Baseline-5")

training <- training[cols_to_retain]
validation <- validation[cols_to_retain]
test <- test[cols_to_retain]

glmmodel <- lm.model(training, validation, model_run)
glmmodel

# Depending on results, either keep or discard columns identified to remove.
check_best_model(err_type_)

#err_type_ <- "MAPE"

```
## Cleanup complete
From the first iteration, we can see that the column removal resulted in a decrease in RMSE.
Rather than just take that, I iterated through the process several times to see if reintroducing the columns
made a change in the process.  In the end the result decreased for all baseline models, and a final baseline of 0.1810351 was achieved.  We will try to improve with Lasso / Ridge / Random Forest / Decision Tree and XG Boost.

**model   alpha  rmse_train**
1	Baseline-1	0	0.1810351	
2	Baseline-2	0	0.1810351	
3	Baseline-3	0	0.1810351	
4	Baseline-4	0	0.1810351	
5	Baseline-5	0	0.1810351	

```{r baseline_5_results, message=FALSE, warning=FALSE}
# Checking the Error values for comparison about whether or not to retain or remove the column suggestions.
print(colResults)

print("Final columns used:")
names(test)

```

### Ridge Regression

For this exercise, we are going to make use of the <a href="https://cran.r-project.org/web/packages/glmnet/index.html">`glmnet`</a> ridge library.

We will also use the same for Lasso, varying alpha values also.

```{r final.predictions.ridge, message=FALSE, warning=FALSE}

lambdas <- 10^seq(-3, 0, by = .05)

train_control_config <- trainControl(method = "repeatedcv", 
                                     number = 5, 
                                     repeats = 1,
                                     returnResamp = "all")

ridge.mod <- caret::train(price ~ ., data = training, 
                          method = "glmnet", 
                          metric = err_type_,
                          trControl=train_control_config,
                          tuneGrid = expand.grid(alpha = 0, lambda = lambdas))

```

The parameter `alpha = 0` means that we want to use the Ridge Regression way of expressing the penalty in regularization. If you replace that by `alpha = 1` then you get Lasso.

#### Evaluation

Plotting the Error for the different lambda values, we can see the impact of this parameter in the model performance.
Small values seem to work better for this dataset.

```{r ridge.plot, message=FALSE, warning=FALSE}

plot(ridge.mod)

```



```{r ridge_coefficients, message=FALSE, warning=FALSE}

plot(ridge.mod$finalModel)

```



```{r ridge.evaluation, message=FALSE, warning=FALSE}

  alpha_ <- 0
  lambda_ <- ridge.mod$bestTune[1,2]
  ridge.mod <- caret::train(price ~ ., data = training, 
                            method = "glmnet", 
                            metric = err_type_,
                            trControl=train_control_config,
                            tuneGrid = expand.grid(alpha = alpha_, lambda = lambda_))


  pred_ <- predict(ridge.mod, validation)
  pred_[is.na(pred_)] <- 0
  real_ <- validation$price

  # Compute Metrics
  rmse_ <- rmse(real_, pred_)
  mape_ <- mape(real_, pred_)
  
  # Error in terms of the mean deviation between the predicted value and the price of the houses
  mae_ <- mae(real_, pred_)
      
  colResults <<- rbind(colResults, data.table(err_id = nrow(colResults)+1, model = "ridge", alpha = alpha_, rmse = rmse_, mae=mae_, mape=mape_));
  
  my_data <- as.data.frame(cbind(predicted=(pred_), observed=(real_)))
  ggplot(my_data, aes(predicted, observed)) +
    geom_point() + geom_smooth(method = "glm") +
    labs(x="Predicted") +
    ggtitle(ggtitle(paste("Ridge :", ': MAPE: ', format(round(mape_, 4), nsmall=4), ' --> Price ERROR:', format(round(mae_, 0), nsmall=0), 
                          ' :: ', sep=''))) +  
    scale_x_continuous(labels = scales::comma) + 
    scale_y_continuous(labels = scales::comma)


# model   alpha     rmse_train
# ridge 0.00000000  1.839320e-01
```

### Rank the variables according to the importance attributed by the model.

```{r plot_ridge, message=FALSE, warning=FALSE}

# Print, plot variable importance
plot(varImp(ridge.mod), top = 20) # 20 most important features

```

### Lasso Regresion

The only thing that changes between Lasso and Ridge is the `alpha` parameter. The remaining part of the exercise is equivalent.

# Final Submission

Based on analysis, we have to decide which cleaning and feature engineering procedures make sense in order to create your final model.

We split the original training data into train and validation to evaluate the candidate models. 
In addition, remember that we also applied a log transformation to the target variable, to revert this transformation you have to use the exp function.


```{r prepare_submission, message=FALSE, warning=FALSE}
# Now for Lasso ... Check some Alpha Values.

alpha = 10^seq(from =-3, to = 0, by = .1) # Goes from 0.01 to 1.

# A .1 improvement on 0.  A 1 No improvement from 0.1.
# Check between 0 and .1 ... and .1 and 1

final.model <- caret::train(price ~ ., data = training,
method = "glmnet",
metric = err_type_,
trControl=train_control_config,
tuneGrid = expand.grid(alpha = 0, lambda = lambdas))

alpha_ = final.model$bestTune[1,1]
lambda_ = final.model$bestTune[1,2]
    
#final.pred <- as.numeric(exp(predict(final.model, validation))-1)
final.pred <- predict(final.model, validation)
final.pred[is.na(final.pred)]

# # Get errors - difference from Train to Val.
predictions_train <- final.pred # Previously taken from log.
train_ColToAnalyse  <-  validation[, "price"];
#train_ColToAnalyse <- as.numeric(exp(train_ColToAnalyse)-1)
errors_train <- predictions_train - train_ColToAnalyse;

# Compute Metrics
rmse_train <- sqrt(mean((predictions_train - train_ColToAnalyse)^2));

lassoResults <- rbind(lassoResults, data.table(err_id = nrow(lassoResults)+1,model = "lasso", alpha = alpha_, lambda = lambda_,rmse_train = rmse_train));

for(alpha_value in alpha){
# Train the model using all the data

    final.model <- caret::train(price ~ ., data = training,
    method = "glmnet",
    metric = err_type_,
    trControl=train_control_config,
    tuneGrid = expand.grid(alpha = alpha_value, lambda = lambdas))
    
    alpha_ = final.model$bestTune[1,1]
    lambda_ = final.model$bestTune[1,2]
    
    # final.pred <- as.numeric(exp(predict(final.model, validation))-1)
    final.pred <- predict(final.model, validation)
    final.pred[is.na(final.pred)]
    
    # # Get errors - difference from Train to Val.
    predictions_train <- final.pred
    train_ColToAnalyse  <-  validation[, "price"];
    #train_ColToAnalyse  <-  as.numeric(exp(train_ColToAnalyse)-1)
    errors_train <- predictions_train - train_ColToAnalyse;
    
    # Compute Metrics
    rmse_train <- sqrt(mean((predictions_train - train_ColToAnalyse)^2));
    
    lassoResults <- rbind(lassoResults, data.table(err_id = nrow(lassoResults)+1,model = "lasso", alpha = alpha_, lambda = lambda_,rmse_train = rmse_train));

}

    alpha_ = lassoResults[lassoResults$rmse_train == min(lassoResults$rmse_train)]$alpha[1]
    lambda_ = lassoResults[lassoResults$rmse_train == min(lassoResults$rmse_train)]$lambda[1]
    
    print (lassoResults[lassoResults$rmse_train == min(lassoResults$rmse_train)])
    
    lassoResults <- NULL
    
    final.model <- caret::train(price ~ ., data = training,
      method = "glmnet",
      metric = err_type_,
      trControl=train_control_config,
      tuneGrid = expand.grid(alpha = alpha_, lambda = lambda_))  
    
    final.pred <- predict(final.model, validation)
    final.pred[is.na(final.pred)]
    

    pred_ <- predict(final.model, validation)
    pred_[is.na(pred_)] <- 0
    real_ <- validation$price
  
    # Compute Metrics
    rmse_ <- rmse(real_, pred_)
    mape_ <- mape(real_, pred_)
    
    # Error in terms of the mean deviation between the predicted value and the price of the houses
    mae_ <- mae(real_, pred_)
        
    colResults <<- rbind(colResults, data.table(err_id = nrow(colResults)+1, model = "lasso", alpha = alpha_, rmse = rmse_, mae=mae_, mape=mape_));
  
    lasso_mod <- final.model

    
```


## Running some final models
Additionally to Baseline Linear models, with Ridge and Lasso, additionally trying Decision Tree, Random Forest and XG Boost

```{r checking_rmseFunctions, message=FALSE, warning=FALSE}

lm_model_rtn<-run_model("glm", "glm")
lm_model <- lm_model_rtn$model
lm_model.pred <- lm_model_rtn$pred
hist(lm_model.pred, main="Histogram of Predictions : Linear Model", xlab = "Predictions")

d_tree_rtn <- run_model("rpart", "rpart")
d_tree <- d_tree_rtn$model
d_tree.pred <- d_tree_rtn$pred
hist(d_tree.pred, main="Histogram of Predictions : Decision Tree", xlab = "Predictions")

rf_tree_rtn <- run_model("rf", "rf")
rf_tree <- rf_tree_rtn$model
rf_tree.pred <- rf_tree_rtn$pred
hist(rf_tree.pred, main="Histogram of Predictions : Random Forest", xlab = "Predictions")

xgb_model_rtn <- run_model("xgboost", "xgboost")
xgb_model <- xgb_model_rtn$model
xgb_model.pred <- xgb_model_rtn$pred
hist(rf_tree.pred, main="Histogram of Predictions : XGBoost", xlab = "Predictions")

    # model   alpha     rmse_train
# ridge   0.00000000  1.839320e-01
# lasso   0.03981072  1.807844e-01
# glm     0.00000000  1.194692e+05
# rpart   0.00000000  1.922892e+05
# rf      0.00000000  1.915833e+05
# xgboost 0.00000000  1.134216e+05

```

## Interpreting Error for all the models run and generating predictions ...


```{r generating_results_files, message=FALSE, warning=FALSE}

best.results <- get_best_model(err_type_);

best.model <- best.results$model[1];
best.mape <- best.results$rmse[1]
best.mae <- best.results$mae[1]
best.rmse <- best.results$rmse[1]

predictions <- 0
val_found=F

if (tolower(best.model)=="glm"){
      # Fit a glm model to the input training data
  predictions = predict(lm_model, test)
  val_found=T
}
if (tolower(best.model)=="rf"){
        # Fit a Random Forest model to the input training data
  predictions = predict(rf_tree$finalModel, test)
  val_found=T 
}
if (tolower(best.model)=="rpart"){
        # Fit a Random Forest model to the input training data
  predictions = predict(d_tree$finalModel, test)
  val_found=T   
}
if (tolower(best.model)=="xgboost"){

  predictions = predict(xgb_model, test)
  val_found=T 
  
}
if((val_found==F) && (substr(tolower(best.model),1, 8)=="baseline")){
  
  base_mod <- get_glm_model(training, validation, "baseline")
  # For Baseline, should always be last model.
  predictions = predict(base_mod, test)
  val_found=T 
  
}
if((val_found==F) && (substr(tolower(best.model),1, 5)=="ridge")){

  predictions = predict(ridge.mod, test)
  val_found=T 
  
}
if((val_found==F) && (substr(tolower(best.model),1, 5)=="lasso")){

  predictions = predict(lasso_mod, test)
  val_found=T 
  
}




if (val_found==T){
  predictions = as.numeric(exp(predictions)-1) 
  final_submission <- data.frame(id = original_test_data$id, price= (predictions))
  colnames(final_submission) <-c("id", "price")
  write.csv(final_submission, file = "AlanCarton_HousePrice_Predictions.csv", row.names = FALSE) 
  
}

best.results

```

## On the discussion board ...


```{r generating_times, message=FALSE, warning=FALSE}

print("Start Time")
print_time(starttime)
print("End Time")
print_time(Sys.time())

```

# Final Results

After data cleaning, columns for the final model were selected as a result of the minimum error increased after their removal.  If the error decreased after removal, the columns were dropped for the final model, otherwise the original columns were retained.

Further models were run with Linear Regression, Decision Tree, Random Forest and XGBoost, and overall the final module used for the predictions and the generation of the final files was chosen based on the lowest error for each model run.  Cross Validation was also used, and the best model was chosen based on the chosen error.

The error code used was set using the err_type_ variable at the top of the code.  Main values were RMSE and MAPE.  MAE was also possible but in all cases the MAE was 0.

Originally the code was run using RMSE, but on review the assignment called for MAPE, so the code was rerun.  Both sets of final details

The models took 26 minutes to run On i5 with 8gb Ram.

## Error Feature Selection and final model selection.

### Lowest results produced 
With feature engineering through RMSE
<!-- model        alpha     rmse      mae   mape -->
<!-- xgboost	  0.05011872	0.1780727	  0	  0.009840251 --> using lowest RMSE 
<!-- xgboost    0.03981072  0.1756873   0   0.009791534 --> using lowest MAPE

Choosing RMSE for feature selection would yield same results, whether using RMSE or MAPE for model selection

### Columns used 
 [1] "price"         "bedrooms"      "bathrooms"     "sqft_living"   "floors"    "waterfront"    "view"        
 [8] "condition"     "grade"         "yr_built"      "yr_renovated"  "zipcode"   "lat"           "sqft_building"
[15] "has_basement"  "increased15"   "unchanged15"  

File generated using RMSE renamed to AlanCarton_HousePrice_Predictions_RMSE.csv

## MAPE Feature Selection and final model selection

### Lowest results produced 
With feature engineering through MAPE
<!-- model        alpha     rmse      mae   mape -->
<!-- xgboost 0.03981072   0.1739514   0     0.009634295 --> using lowest RMSE
<!-- xgboost 0.03981072   0.1739514   0     0.009634295 --> using lowest MAPE

### Columns used
 [1] "price"           "bedrooms"        "bathrooms"       "sqft_living"     "sqft_lot"        "floors"         
 [7] "waterfront"      "view"            "condition"       "grade"           "sqft_above"      "sqft_basement"  
[13] "yr_built"        "yr_renovated"    "zipcode"         "lat"             "long"            "sqft_living15"  
[19] "sqft_lot15"      "TotalFlArea"     "sqft_building15" "sqft_building"  

File generated using RMSE renamed to **AlanCarton_HousePrice_Predictions_MAPE.csv **

## Best model results obtained by selecting the columns using MAPE / and either MAPE or RMSE for model selection, and All Columns were used including additional features.  MAPE for final results 0.0096 MAPE or 0.1739 RMSE

In both instances XGBoost was the best model, the difference for the final dataset being as a result of the column selection as a result of MAPE or RMSE, with the best combination being MAPE selection, and either RMSE or MAPE error checking producing an RMSE of 0.1739514 and MAPE of 0.009634295
